Monitoring forest health using hyperspectral imagery: Does feature selection improve the performance of machine-learning techniques?

### 2021-11-15 docs/00-manuscripts/mdpi/manuscript.tex
revising discussion (1)

The best aggregated performance of this study (SVM + \enquote{Info Gain} filter, RMSE~27.915 p.p.) has to be seen in the light of model overfitting (see \autoref{subsec:perf-plot-char}).
Leaving out the performance on Laukiz2 when aggregating results, the mean RMSE would be around 19 p.p.
However, leaving out a single plot would also change the prediction results for the other plots because the observations from Laukiz2 would not be available for model training.
Due to the apparent presence of model overfitting in this study, it can be postulated that more training data representing a greater variety of situations is needed.
A model can only make robust predictions if it has learned relationships across the whole range of the response.
Hence, care should be taken when predicting to the landscape scale using models fitted on this dataset due to their lack of generalizability caused by the limitations of the available training data.
However, when inspecting the fold\-/level performances, it can be concluded that the models performed reasonably well predicting defoliation greater than 50\% but failed for lower levels.
This applied to all learners of this study.
The overall performance of all learners achieved in this study can be classified as ``poor'' given that only the SVM learner was able to outperform the featureless learner (\autoref{tab:best-learner-perf}).

In addition, data quality issues (\autoref{subsec:data-quality}) might have an influence on model performances.
These include the timing of the acquisition of the hyperspectral data (late phenological phase), field measurement errors when surveying defoliation, the influence of background reflectance (e.g., soil reflectance) and the possible positional offset of measured GPS coordinates of trees.

@@

An interesting finding is the strength of the SVM algorithm when comparing its predictive performance to its competitors (\autoref{tab:best-learner-perf}).
These cluster around a performance of 31 p.p. while SVM is able to score about three p.p. better than all other methods.
However, we refrain from comparing these results (both relatively and absolute) to other studies since many study design points have an influence on the final result (optimization strategy, data characteristics, feature selection methods, etc.).

@@

One objective of this study was whether expert-based or data-driven feature engineering has a positive influence on model performance.
With respect to \autoref{fig:perf-result}, no overall positive or negative trend was found for all models that related to specific feature sets.
The performance of RF and XGBoost on the VI feature set was around four to six p.p. lower than on others.
One reason could be the lack of coverage in the wavelength area between 810~nm and 1000~nm (\autoref{fig:fi-permut-vi-hr}).
In addition, for all learners but SVM a better performance was observed when NRI indices were included in the feature set (i.e., NRI-VI, HR-NRI, HR-NRI-VI).

@@

Arbitrary seeming values of selected feature percentages in \autoref{tab:tune-perc-sel-features} emerge from the specific hyperparameter setting which got elected as the best one during the tuning step.
The specific performance value to benchmark against during tuning is defined by the internal surrogate learner of the MBO tuning method and can be any numeric value between 0 and 1 (interpreted as a\ percentage of available features) for the hyperparameter that controls the number of features.

% filters always reduce computation time
Realizing early during hyperparameter optimization that only few features are needed to reach adequate performances can reduce the overall computational runtime substantially.
Hence, regardless of the potential advantage of using filters for increased predictive performance, these can have a strong positive effect on runtime, especially on models which make use of hyperparameters that depend on the available number of features, such as RF $m_\textrm{try}$.

Ultimately, the results of \autoref{tab:tune-perc-sel-features} should be taken with care as they rely on single model-filter combinations and are subject to random variation.
More in-depth research is needed to investigate the effect of filters on other criteria than performance (such as runtime), leading to a multi-criterion optimization problem.

>>>>

The best overall performance achieved in this study (SVM with the \enquote{Info Gain} filter, RMSE~27.915 p.p.) has to be seen in the light of model overfitting (see \autoref{subsec:perf-plot-char}).
Leaving out the performance on Laukiz2 when aggregating results, the mean RMSE would be around 19 p.p.
However, leaving out a single plot would also change the prediction results for the other plots because the observations from Laukiz2 would not be available for model training.
Due to the apparent presence of model overfitting in this study, it is suggested that more training data representing a greater variety of situations is needed.
A model can only make robust predictions if it has learned relationships across the whole range of the response.
Hence, care should be taken when predicting on the landscape scale using models fitted on this dataset due to their lack of generalizability caused by the limitations of the available training data.
However, when inspecting the fold\-/level performances, it can be concluded that the models performed reasonably well predicting defoliation greater than 50\%, but they failed for lower levels.
This applied to all learners of this study.
In this study, the overall performance across all learners can be classified as ``poor'' given that only the SVM learner was able to substantially outperform the featureless learner (\autoref{tab:best-learner-perf}).
It is worth noting that data quality issues may have affecting model performances, as discussed below in detail  (\autoref{subsec:data-quality}).

@@

An interesting finding is the strength of the SVM algorithm when comparing its predictive performance to its competitors (\autoref{tab:best-learner-perf}).
These cluster around a performance of 31 p.p. while SVM scored about three p.p. better than all other methods.
However, we refrain from comparing these results (both relatively and absolute) to other studies since many study design points have an influence on the final result (optimization strategy, data characteristics, feature selection methods, etc.).

@@

One objective of this study was to determine whether expert-based and data-driven feature engineering have a positive influence on model performance.
With respect to \autoref{fig:perf-result}, no overall positive or negative pattern related to specific feature sets was found that would be valid across all models.
The performance of RF and XGBoost on the VI feature set was around four to six p.p. lower than on others.
One reason could be the lack of coverage in the wavelength range between 810~nm and 1000~nm (\autoref{fig:fi-permut-vi-hr}).
In addition, for all learners but SVM a better performance was observed when NRI indices were included in the feature set (i.e., NRI-VI, HR-NRI, HR-NRI-VI).

@@

There was also no clear pattern in the percentage of features selected based on hyperparameter tuning (\autoref{tab:tune-perc-sel-features}).
The specific performance value to benchmark against during tuning is defined by the internal surrogate learner of the MBO tuning method and can be any numeric value between 0 and 1 (interpreted as a\ percentage of available features) for the hyperparameter that controls the number of features.
% Nevertheless, considering SVM's relative success, the use of at least a few hundred features from the combined feature set appears to be beneficial, or at least not harmful when the model's built-in regularization is capable of dealing with the resulting high-dimensional situation.

% filters always reduce computation time
Realizing early during hyperparameter optimization that only few features are needed to reach adequate performances can reduce the overall computational runtime substantially.
Hence, regardless of the potential advantage of using filters for increased predictive performance, these can have a strong positive effect on runtime, especially of models making use of hyperparameters that depend on the available number of features, such as RF with $m_\textrm{try}$.

Ultimately, the results of \autoref{tab:tune-perc-sel-features} should be taken with care as they rely on single model-filter combinations and are subject to random variation.
More in-depth research is needed to investigate the effect of filters on other criteria than performance (such as runtime), leading to a multi-criteria optimization problem.
